---
title: "purify"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{purify}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

Welcome to `purify'! This vignette will outline the main functions and offer many valuable simulations for these funtions. Real data examples can be found in other vignettes.

`Purify' comes from its purpose of permutation tests and statification for data. It can be used to purify data and mitigate confounding effects, particularly when looking at strata which are much smaller than other strata.

The data can be found on GitHub and can be loaded as:
```{r setup}
library(purify)
```

## Example from paper to include

The following example demonstrates the use of `Purify` with simulated data, applying a mean squared error (MSE) calculation over 1,000 resamples using the simple method:
```{r}
# Simulate data
set.seed(123)
n <- 100
data <- data.frame(
  output = rnorm(n),
  predictor1 = rnorm(n),
  predictor2 = rnorm(n)
)

# Define a custom function to calculate MSE
mse_function <- function(data) {
  mean((data$output - predict(lm(output ~ ., data = data)))^2)
}
coef_function <- function(data) {
  coef(lm(output ~ ., data = data))
}

summ_function <- function(data) {
  coef(summary(lm(output ~ ., data = data)))
}

# Perform resampling with the 'simple' method
results <- resample_function(data = data, fn = mse_function, 
                             M = 1000, method = 'simple')
results$estimates
results_coef <- resample_function(data = data, fn = coef_function, 
                                  M = 1000, method = "simple")
results_coef$estimates
results_summ <- resample_function(data = data, fn = summ_function, 
                                  M = 1000, method = "simple")
results_summ$estimates
```

Model coefficients can also be examined under permutation; see documentation. However, if a use case is too complex for the implemented `resample_function()`, the function `resample()` can be used to directly return the resampled data.


# Using resample_function()

In this vignette, we explore the use of this package with simulated data. The function \code{resample_function()} has four options, 'simple', 'stratify', 'sliding', 
and 'segment'. We follow these for this vignette.


## Simple Permutation

Consider a simple data set as generated below.
```{r simple_generate}
set.seed(12345)
n <- 100

data_simple <- data.frame('y'=rep(NA,n),
                          'x'=rnorm(n),
                          'z'=rexp(n))
outliers <- runif(3,1,n)
data_simple[outliers,'x'] <- 25*data_simple[outliers,'x']
data_simple[outliers,'z'] <- 10*data_simple[outliers,'z']
errors <- rnorm(n,0,5)
errors[outliers] <- 10*errors[outliers]
data_simple$y <- 1 + 2*data_simple$x - 3*data_simple$z + data_simple$x*data_simple$z + errors

```

Suppose we are interested in the regression $$y\sim \beta_0+\beta_1x+\beta_2z+\beta_3xz$$. This can be codified as follows, noting that the first value of the output data.frame is the estimates.
```{r simple_function}
simple_function <- function(X){
  reg <- lm(y~x*z, data=X)
  
  coef(summary(reg))
}
```

The model on the original data is computed.
```{R simple_regression}
set.seed(12345)
simple_function(data_simple)
```
This simulation returns value which are similar but are definitely different from the true values
This can be resampled as follows.
```{r simple_resample_function}
set.seed(12345)
res <- resample_function(data_simple, simple_function)
res$estimates
```
The resampling returns better estimates.


## Statified Permutation (TODO::UPDATE)

Consider a stratified data set as generated below.
```{r stratify_generate}
set.seed(1234)
n <- 500

data_strat <- data.frame('y'=rep(NA,n),
                         'x'=NA,
                         'z'=sample(c('A','B','C'),size=n,replace=TRUE,prob=c(0.65,0.3,0.05)),
                         'w'= rnorm(n,75,50))
data_strat[data_strat$z=='A',"x"] <- rnorm(length(data_strat[data_strat$z=='A',"x"]),40,15)
data_strat[data_strat$z=='B',"x"] <- rnorm(length(data_strat[data_strat$z=='B',"x"]),60,25)
data_strat[data_strat$z=='C',"x"] <- rnorm(length(data_strat[data_strat$z=='C',"x"]),100,25)

data_strat$y <- 10 + 0.1*data_strat$x + rnorm(n,sd=10)

```

We see major in balance is the groups
```{r, stratify_imbalance}
boxplot_strata(data_strat,'z')
```

Suppose we are interested in the following regression model.
```{r stratify_function}
strat_function <- function(X){
  reg <- lm(y~x+z+w, data=X)
  
  coef(summary(reg))
}

```

The model on the original data can be computed.
```{r strat_regression}
set.seed(1234)
strat_function(data_strat)
```
This simulation returns value which are far different from the true values,
This can be resampled as follows.
```{r strat_resample_function}
set.seed(1234)
res <- resample_function(data = data_strat, fn = strat_function, method='stratify',
                         strata='z', size=200)
res$estimates
```
The resampling returns better estimates.


## Dependent (Sliding / Segment) Permutation

Consider a dependent data set as generated below.
```{r dependent_generate}
set.seed(1234)
n <- 500

burnin <- 500
data_ts <- data.frame('Y'=rep(NA,n+burnin),
                      'X'=rnorm(n+burnin))
for(i in 1:(n+burnin)){
  if(i>5){
    data_ts$Y[i] <- data_ts$X[i] + sum(seq(5,3,-0.5)/10*data_ts$X[i-1:5]) + rnorm(1)
  }else{
    data_ts$Y[i] <- data_ts$X[i] + rnorm(1)
  }
}
data_ts <- data_ts[-(1:burnin),]

```

We could resample as follows.
```{r dependent_resample}
res <- resample(data = data_ts,method='sliding',blockSize = 5)
res1 <- resample(data_ts,method='segment',blockSize = 5)
```
