---
title: "purify"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{purify}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

Welcome to `purify'! This vignette will outline the main functions and offer many valuable simulations for these funtions. Real data examples can be found in other vignettes.

`Purify' comes from its purpose of permutation tests and statification for data. It can be used to purify data and mitigate confounding effects, particularly when looking at strata which are much smaller than other strata.

The data can be found on GitHub and can be loaded as:
```{r setup}
library(purify)
```

# Simulations 

Let the data be generated as follows.

```{r stratify_generate}
set.seed(1234)
n <- 500

data_strat <- data.frame(
  "y" = rep(NA, n),
  "x" = NA,
  "z" = sample(c("A", "B", "C"), size = n, replace = TRUE, prob = c(0.65, 0.3, 0.05)),
  "w" = rnorm(n, 75, 50)
)
data_strat[data_strat$z == "A", "x"] <- rnorm(length(data_strat[data_strat$z == "A", "x"]), 40, 15)
data_strat[data_strat$z == "B", "x"] <- rnorm(length(data_strat[data_strat$z == "B", "x"]), 60, 25)
data_strat[data_strat$z == "C", "x"] <- rnorm(length(data_strat[data_strat$z == "C", "x"]), 100, 25)

data_strat$y <- 10 + 0.1 * data_strat$x + rnorm(n, sd = 10)
```

We see a major inbalance is the groups.
```{r, stratify_imbalance}
plot_strata_bar(data_strat$z)
plot_strata_box(data_strat[, c("z", "x")])
```

Suppose we are interested in the following regression model.
```{r stratify_function}
strat_function <- function(X) {
  reg <- lm(y ~ x + z + w, data = X)

  coef(summary(reg))
}
```

The model on the original and resampled data can be computed.
```{r strat_regression}
set.seed(1234)
orig_coefs <- strat_function(data_strat)

set.seed(1234)
res <- resample(
  data = data_strat, fn = strat_function,
  strata = "z", sizes = "mean"
)
```

The results show resampling improved the estimates.
```{r comparision}
orig_coefs
summarize_resample(res)
```

# Titanic Example

From a subset of titanic data, it may be natural to try to estimate the survival 
of a passenger.

```{r titanic}
# Define a custom function to calculate MSE
mse_function <- function(data) {
  pred <- as.numeric(predict(
    glm(as.factor(Survived) ~ .,
      data = data,
      family = binomial(link = "logit")
    ),
    type = "response"
  ) > 0.5)
  mean(as.numeric(data$Survived != pred))
}

# Simple resampling
set.seed(1234)
results <- resample(
  data = titanic, fn = mse_function,
  M = 1000, strata = "Pclass"
)
```

Resampling the data lead to lower mean squared error estimates.

```{r errors}
# Does CV on each sample
#   Note, MSE uses all data while resample will be missing the CV part
#   See cats vignette for modifications
mse_function1 <- function(data) {
  cv <- cross_validation(
    data = data,
    pred_fn = function(data, nd) {
      as.numeric(predict(
        glm(as.factor(Survived) ~ .,
          data = data,
          family = binomial(link = "logit")
        ),
        type = "response", newdata = nd
      ))
    },
    error_fn = function(data, pred) {
      mean(as.numeric(data$Survived != pred))
    }
  )
  mod <- glm(as.factor(Survived) ~ .,
    data = data,
    family = binomial(link = "logit")
  )

  c(cv[[1]], as.numeric(coef(summary(mod))[, 1]))

  # mod <- lm(Hwt ~ ., data=data)
  # pred <- as.numeric(predict( mod , new_data=cv_data) )
  # c(mean((data$Hwt - pred)^2), as.numeric(coef(summary(mod))[,1]) )
}

set.seed(1234)
results <- resample(
  data = titanic, fn = mse_function,
  M = 1000, strata = "Pclass"
)

mse_function(titanic)
summarize_resample(results)
```

# German Credit Information

Below is a simple analysis of German credit data.

```{r credit}
set.seed(1234)
# Define a custom function to calculate MSE
mse_function <- function(data) {
  pred <- unique(data$Class)[
    as.numeric(
      predict(
        glm(as.factor(Class) ~ .,
          data = data,
          family = binomial(link = "logit")
        ),
        type = "response"
      ) > 0.5
    ) + 1
  ]

  sum(as.numeric(data$Class != pred))
}

# Perform resampling with the 'simple' method
#   Note, MSE uses all data while resample will be missing the CV part
#   See cats vignette for modifications
results <- resample(
  data = credit, fn = mse_function,
  M = 2000, strata = "Job"
)
summarize_resample(results)
mse_function(credit)
```

