---
title: "Simulations"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Simulations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Overview and Setup

```{r setup}
library(purify)
```

In this vignette, we explore the use of this package with simulated data. The function \code{resample_function()} has four options, 'simple', 'stratify', 'sliding', 
and 'segment'. We follow these for this vignette.


# Simple Permutation

Consider a simple data set as generated below.
```{r simple_generate}
set.seed(12345)
n <- 100

data_simple <- data.frame('y'=rep(NA,n),
                          'x'=rnorm(n),
                          'z'=rexp(n))
outliers <- runif(3,1,n)
data_simple[outliers,'x'] <- 25*data_simple[outliers,'x']
data_simple[outliers,'z'] <- 10*data_simple[outliers,'z']
errors <- rnorm(n,0,5)
errors[outliers] <- 10*errors[outliers]
data_simple$y <- 1 + 2*data_simple$x - 3*data_simple$z + data_simple$x*data_simple$z + errors

```

Suppose we are interested in the regression $$y\sim \beta_0+\beta_1x+\beta_2z+\beta_3xz$$. This can be codified as follows, noting that the first value of the output data.frame is the estimates.
```{r simple_function}
simple_function <- function(X){
  reg <- lm(y~x*z, data=X)
  
  coef(summary(reg))
}
```

The model on the original data is computed.
```{R simple_regression}
set.seed(12345)
simple_function(data_simple)
```
This simulation returns value which are similar but are definitely different from the true values
This can be resampled as follows.
```{r simple_resample_function}
set.seed(12345)
res <- resample_function(data_simple, simple_function)
res$estimates
```
The resampling returns better estimates.


# Statified Permutation (TODO::UPDATE)

Consider a stratified data set as generated below.
```{r stratify_generate}
set.seed(1234)
n <- 500

data_strat <- data.frame('y'=rep(NA,n),
                         'x'=NA,
                         'z'=sample(c('A','B','C'),size=n,replace=TRUE,prob=c(0.65,0.3,0.05)),
                         'w'= rnorm(n,75,50))
data_strat[data_strat$z=='A',"x"] <- rnorm(length(data_strat[data_strat$z=='A',"x"]),40,15)
data_strat[data_strat$z=='B',"x"] <- rnorm(length(data_strat[data_strat$z=='B',"x"]),60,25)
data_strat[data_strat$z=='C',"x"] <- rnorm(length(data_strat[data_strat$z=='C',"x"]),100,25)

data_strat$y <- 10 + 0.1*data_strat$x + rnorm(n,sd=10)

```

We see major in balance is the groups
```{r, stratify_imbalance}
boxplot_strata(data_strat,'z')
```

Suppose we are interested in the following regression model.
```{r stratify_function}
strat_function <- function(X){
  reg <- lm(y~x+z+w, data=X)
  
  coef(summary(reg))
}

```

The model on the original data can be computed.
```{r strat_regression}
set.seed(1234)
strat_function(data_strat)
```
This simulation returns value which are far different from the true values,
This can be resampled as follows.
```{r strat_resample_function}
set.seed(1234)
res <- resample_function(data = data_strat, fn = strat_function, method='stratify',
                         strata='z', size=200)
res$estimates
```
The resampling returns better estimates.


# Dependent (Sliding / Segment) Permutation

Consider a dependent data set as generated below.
```{r dependent_generate}
set.seed(1234)
n <- 500

burnin <- 500
data_ts <- data.frame('Y'=rep(NA,n+burnin),
                      'X'=rnorm(n+burnin))
for(i in 1:(n+burnin)){
  if(i>5){
    data_ts$Y[i] <- data_ts$X[i] + sum(seq(5,3,-0.5)/10*data_ts$X[i-1:5]) + rnorm(1)
  }else{
    data_ts$Y[i] <- data_ts$X[i] + rnorm(1)
  }
}
data_ts <- data_ts[-(1:burnin),]

```

We could resample as follows.
```{r dependent_resample}
res <- resample(data = data_ts,method='sliding',blockSize = 5)
res1 <- resample(data_ts,method='segment',blockSize = 5)
```
