---
title: "cats"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Overview

This article examines data on cats, includes as *cats* and *subcats* in the package. The goal of this analysis is to see the effect sex has on the body and heart weight measures. This analysis goes further than that in the purify article by adding cross-validation to the summary function. This comes at the cost of computational time.
```{r setup}
library(purify)
library(ggplot2)

summ_function <- function(data) {
  cv <- cross_validation(
    data = data,
    pred_fn = function(data, nd) {
      as.numeric(predict(lm(Hwt ~ ., data = data), newdata = nd))
    },
    error_fn = function(true, est) {
      mean((true$Hwt - est)^2)
    }
  )
  mod <- lm(Hwt ~ ., data = data)
  c(cv[[1]], as.numeric(coef(summary(mod))[, 1]))
}
```

# Full Data

We first are interested in the coefficients when estimating the body weight using sex and heart weight, stratified on sex since there are far less female cats observed.
```{r strata_plots}
plot_strata_bar(cats$Sex)
plot_strata_box(cats[, c("Sex", "Hwt")])
plot_strata_box(cats[, c("Sex", "Bwt")])

ggplot() +
  geom_point(aes(x = Bwt, y = Hwt, col = Sex, shape = Sex), data = cats, size = 5) +
  theme_bw() +
  theme(
    axis.title = element_text(size = 30),
    axis.text = element_text(size = 26),
    legend.position = c(.2, .8),
    legend.title = element_blank(),
    legend.text = element_text(size = 26)
  ) +
  scale_color_discrete(labels = c("Female", "Male")) +
  scale_shape_manual(
    labels = c("Female", "Male"),
    values = c(16, 3)
  ) +
  xlab("Body weight (kg)") +
  ylab("Heart weight (g)")
```

Now modeling the data:
```{r full_estimate}
set.seed(1234)
summ_function(cats)

# Perform resampling (with CV included so a bit time consuming)
results <- resample(
  data = cats, fn = summ_function, M = 500,
  strata = "Sex", sizes = mean
)
summarize_resample(results)
```

In both methods, the sex was insignificant and body weight was important.

# Subset Data

We are now interested in the coefficients when estimating the body weight using sex and heart weight, stratified on sex since there are far less female cats observed, with a smaller data set. In practice it is hard to know the exact effect of data sizes, so it is worth examining what may happen. We subset the data such that sex will now be more important
```{r substrata_plots}
plot_strata_bar(subcats$Sex)
```

Now modeling the data:
```{r subestimate}
# Original model
set.seed(1234)
tmp <- lm(Hwt ~ ., data = subcats)
coef(summary(tmp))
confint(tmp)

## CV for both
set.seed(1234)
results <- resample(
  data = subcats, fn = summ_function,
  M = 500, strata = "Sex"
)

cross_validation(
  data = subcats,
  pred_fn = function(data, nd) {
    as.numeric(predict(lm(Hwt ~ ., data = data), newdata = nd))
  },
  error_fn = function(true, est) {
    mean((true$Hwt - est)^2)
  }
)
summarize_resample(results)
```

In examining the data, we can observe that sex is important, but is hard to see due to the low samples. Traditional linear models miss this; however, a resampled method can capture the importance of sex without losing the effects of body weight. The resampled
method also improves estimation in terms of mean squared error.
